---
title: "Regeneration_3population"
author: "Tyler Harvey"
date: "4/10/2018"
output: html_document
---
## Load programs [do I need to point to gene_exp.diff??]

- Point to Cuffdiff output for 3 populations from ECESCX
```{r}
.libPaths( "/mnt/sequence/harvey/git/tendon/Dual_GFP_Tom_regeneration_diffout" )
.libPaths()
```

## Load programs and configure [need to load CummeRbund package and others]


```{r}
source("https://bioconductor.org/biocLite.R")
biocLite("cummeRbund")

library( "cummeRbund")
library( "dplyr")
library( "tidyr")

options( "width"=100 )
```

## Load data



```{r}
cuff <- readCufflinks( "results")
cuff
```


```{r}
runInfo( cuff)[1:2,]
```

##Diagnostic plots

```{r}
MDSplot( genes( cuff), replicates=TRUE )
MDSplot( genes( cuff ), "Dual", "GFPinj" )
MDSplot( genes( cuff ), "Dual", "Tominj" )
MDSplot( genes( cuff ), "GFPinj", "Tominj" )
```

##Check positive/negative controls GOIs
```{r}
goi <- c( "Tppp3","Pdgfra","Scx","Mkx","Bgn","Ctgf","Dcn","Fn1","Fmod","Tnc","Tnmd")
expressionPlot( getGenes(cuff,goi), replicates=TRUE, showStatus=FALSE) +
  facet_wrap("tracking_id",scales="free_y")+
  theme( legend.position="none")
```

```{r}
diffData( genes( cuff ) ) %>%
  filter( gene_id %in% goi,
          !(sample_1 == "Dual" & sample_2 == "GFPinj" ) ) %>%
  select( -test_stat, -p_value ) %>%
  arrange( gene_id )
```

##First, only consider genes with FPKM > 5 in any condition [want to be >5 in at least one, not necessarily both]
```{r}
goi_gt5 <- diffData( genes( cuff ) ) %>%
  filter( !( sample_1 == "Dual" & sample_2 == "GFPinj" ) ) %>%
  group_by( gene_id ) %>%
  summarize( value_max=max( value_1, value_2 ) ) %>%
  filter( value_max > 5 ) %>%
  pull( gene_id )
length( goi_gt5 )
```



##Cluster using Jensen-Shannon distance and partitioning around medoids
```{r}
a <- getGenes( cuff, df$gene_id )
clusters <- c(5,8)
for ( i in clusters ) {
  ic <- csCluster( a, k=i )
  df <- cbind(df, ic$clustering[ match( df$gene_id, names( ic$clustering ) ) ] )
  
  p <- csClusterPlot( ic ) +
    facet_wrap( "cluster" ) +
    geom_hline( yintercept=2, color="red" )+
    annotate( "text", label=paste( ic$clusinfo[ ,"size" ], "genes" ), x=2, y=4.25) +
    theme(legend.position="none" )
  print( p )
}

names( df ) <- c("gene_id", paste( "cluster", clusters, sep=""))
filter(df, gene_id %in% goi)
write.csv( df, file="clusters.both.csv", row.names = FALSE)
```

##Document program versions
```{r}
sessionInfo()
```
