---
title : "Tyler Harvey - TH1"
author: "Frederick Tan"
date  : "March 29, 2018"
output: html_notebook
---

## Load programs

- Point to 10x library

```{r}
.libPaths( "/mnt/sequence/10x/R/3.4.3" )
.libPaths()
```

- Load Cell Ranger R Kit, etc.

```{r message=FALSE}
library( "cellrangerRkit" )
library( "Seurat" )
library( "monocle" )

library( "stringr" )
library( "dplyr" )
```


## Load data

```{r}
TH1.gbm <- load_cellranger_matrix( "/mnt/sequence/10x/TH1_count" )
TH1.analysis <- load_cellranger_analysis_results( "/mnt/sequence/10x/TH1_count" )

TH1.data <- Read10X( "/mnt/sequence/10x/TH1_count/outs/filtered_gene_bc_matrices/mm10" )
TH1.seurat <- CreateSeuratObject( TH1.data, project="TH1" )

TH1 <- importCDS( TH1.seurat )
```


## Visualize Cell Ranger results

- clusters
    - 2: sheath ... Tppp3, some Pdgfra
        - dual vs tom
    - 3: FAP ... Pdgfra
        - gfp
    - 4: tenocyte ... Scx Mkx

```{r}
tsne_proj <- TH1.analysis$tsne
visualize_clusters( TH1.analysis$clustering$graphclust$Cluster, 
                    tsne_proj[ c( "TSNE.1", "TSNE.2" ) ],
                    marker_size=1.0 )
```

- marker genes

```{r}
use_genes <- get_nonzero_genes( TH1.gbm )
gbm_bcnorm <- normalize_barcode_sums_to_median( TH1.gbm[ use_genes, ] )
gbm_log <- log_gene_bc_matrix( gbm_bcnorm, base=10 )

goi <- c( "Tppp3", "Scx", "Mkx", "Pdgfra", "Acta2" )
visualize_gene_markers( gbm_log, goi, tsne_proj[ c( "TSNE.1", "TSNE.2" ) ] )
```


## Annotate Monocle cells with Cell Ranger clusters

```{r}
table( rownames( pData( TH1 ) ) == str_sub( TH1.analysis$clustering$graphclust$Barcode, 1, -3 ) )
pData( TH1 )$Cluster10x = TH1.analysis$clustering$graphclust$Cluster
pData( TH1 )
```

#### Tabulate cells with 1 or more marker gene transcripts

```{r}
goi <- c( "Tppp3", "Scx", "Mkx", "Pdgfra", "Acta2" )
for (i in goi ) {
  pData( TH1 )[ , i ] <- exprs( TH1 )[ i, ] >= 1
}

pData( TH1 ) %>%
  mutate ( Acta2Tppp3 = Acta2 & Tppp3) %>%
  mutate ( Tppp3Acta2 = Tppp3 & Acta2) %>%
  group_by( Cluster10x ) %>%
  summarize( n=n(),
             Tppp3=sum( Tppp3==TRUE ),
             Scx=sum( Scx==TRUE ),
             Mkx=sum( Mkx==TRUE ),
             Pdgfra=sum( Pdgfra==TRUE ),
             Acta2=sum( Acta2==TRUE),
             Acta2Tppp3=sum (Acta2Tppp3==TRUE),
             Tppp3Acta2=sum (Tppp3Acta2==TRUE))
```

## Re-analyze sheath and tenocytes

#### Subset cells

```{r}
coi <- pData( TH1 )$Cluster10x %in% c( 2, 3, 4 )
TH1_234 <- TH1[ , coi ]
```

#### Cluster cells

```{r}
TH1_234 <- estimateSizeFactors( TH1_234 )
TH1_234 <- estimateDispersions( TH1_234 )

TH1_234 <- detectGenes( TH1_234, min_expr=0.1 )
fData( TH1_234 )$use_for_ordering <- fData( TH1_234 )$num_cells_expressed > 0.05 * ncol( TH1_234 )

TH1_234 <- reduceDimension( TH1_234, reduction_method='tSNE' )
TH1_234 <- clusterCells( TH1_234 )

table( pData( TH1_234 )[ , c( "Cluster", "Cluster10x" ) ] )
plot_cell_clusters( TH1_234, color_by="Cluster" )
```

#### Construct trajectory

- ~20 minutes
- VGAM not found if `cores=8`

```{r eval=FALSE}
TH1_24_expressed_genes <- row.names( subset( fData( TH1_24 ), num_cells_expressed >= 10 ) )
TH1_24_clustering_DEG_genes <- differentialGeneTest( TH1_24[ TH1_24_expressed_genes, ], fullModelFormulaStr="~Cluster" )
save( TH1_24_clustering_DEG_genes, file="TH1_24_clustering_DEG_genes.Rda" )
```

```{r}
load( "TH1_24_clustering_DEG_genes.Rda" )

TH1_24_ordering_genes <- row.names( TH1_24_clustering_DEG_genes )[ order( TH1_24_clustering_DEG_genes$qval ) ][ 1:1000 ]
TH1_24 <- setOrderingFilter( TH1_24, ordering_genes=TH1_24_ordering_genes )

TH1_24 <- reduceDimension( TH1_24, method="DDRTree" )
TH1_24 <- orderCells( TH1_24 )

plot_cell_trajectory( TH1_24, color_by="State" )
plot_genes_jitter( TH1_24[ goi, ], grouping="State" )
plot_genes_in_pseudotime( TH1_24[ goi, ], color_by="State" )
```

#### Tabulate markers in states

```{r}
pData( TH1_24 ) %>%
  group_by( State ) %>%
  summarize( n=n(),
             Tppp3=sum( Tppp3==TRUE ),
             Scx=sum( Scx==TRUE ),
             Mkx=sum( Mkx==TRUE ),
             Pdgfra=sum( Pdgfra==TRUE ) )
```





